### 2012-12-20 paper.tex
edits part 1

%%% template.tex
%%%
%%% This LaTeX source document can be used as the basis for your technical
%%% paper or abstract. Intentionally stripped of annotation, the parameters
%%% and commands should be adjusted for your particular paper - title, 
%%% author, article DOI, etc.
%%% The accompanying ``template.annotated.tex'' provides copious annotation
%%% for the commands and parameters found in the source document. (The code
%%% is identical in ``template.tex'' and ``template.annotated.tex.'')

\documentclass[annual]{acmsiggraph}

\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{1111111.2222222}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

\usepackage{caption}
\usepackage{subcaption}
\usepackage[english]{babel}
\title{Spec2Fab: A  Graph-Based Framework for Fabricating Fabrication Algorithms}

\author{}
\pdfauthor{}

\keywords{}

\newcommand{\note}[1]{\marginpar{\LARGE $\spadesuit$}
			$\spadesuit$ {\bf #1} $\spadesuit$}

\begin{document}

 \teaser{
   \begin{subfigure}[b]{0.3\textwidth}
	\centering
    	\includegraphics[width=0.6\textwidth]{figure/bunny_orig.png}
      	\caption{}
        \label{teaser:original}
        \end{subfigure}
        ~ 
        \begin{subfigure}[b]{0.18\textwidth}
                \centering
    	\includegraphics[width=\textwidth]{figure/bunny_print1.jpg}
                \caption{}
                \label{teaser:ours}
        \end{subfigure}
        ~
        \begin{subfigure}[b]{0.2\textwidth}
                \centering
    	\includegraphics[width=\textwidth]{figure/bunny_print2.jpg}
                \caption{}
                \label{teaser:makerbot}
        \end{subfigure}
        ~
        \begin{subfigure}[b]{0.18\textwidth}
                \centering
    	\includegraphics[width=\textwidth]{figure/bunny_print3.jpg}
                \caption{}
                \label{teaser:objet}
        \end{subfigure}
        \caption{A textured mesh printed with different types of 3D printers.
        	(a)original model.(b)printed with our printer.(c)makerbot.
        	(d)objet}\label{teaser}
	\label{fig:texture}
 }

\maketitle

\begin{abstract}

\end{abstract}

\keywordlist

\TOGlinkslist

\copyrightspace

\note{Teaser should be all different results produced using spec2Fab, texture, subsurface scattering, caustics etc...}

\section{Introduction}
Computational fabrication couples new digital manufacturing techniques with novel numerical algorithms to produce objects with remarkable physical characteristics. These methods are of great interest in computer graphics for two reasons: First, they provide us with  new class of output devices which can render into  real-life and second, they yield valuable insight into the accuracy of our numerical models of the world.

In this paper we focus on a fundamental problem in producing any such custom object:  optimal material assignment. Given an object one must find a spatially varying material distribution that satisfies some given criteria. The algorithms that solve these problems are often complicated to develop, implement, test and modify. We provide a new graph-based framework which decomposes such problems into a small set of constituent components alleviating the above issues as well as facilitating the development and testing of novel algorithms.

\section{Previous Work}
\subsection{Simulation of the Appearance of Ink Combinations}
\cite{Matusik:2009} built a system
for the measurement of Bidirectional Reflectance Distribution Functions (BRDFs) and were able
to acquire BRDFs for both single-layer and multi-layer ink printouts (e.g., an ink covered by a
varnish layer). They then acquired a database of many different combinations of ink layers (up to
seven layer combinations were considered). However, they did not develop a method to simulate
the appearance of arbitrary ink layers. They were able to interpolate the appearance by spatially
multiplexing small tiles of different base combinations. By linearity of light transport, the simulation
of the combined appearance was reduced to linear combinations of the base BRDFs. 
simulation method for the appearance of translucent materials \cite{Hasan:2010}. 
They
built a measurement system for estimating multiple-scattering in homogeneous material slabs. They
measured both transmission and refection for a set of base materials used by the OBJET Connex
multi-material printer. 
Using these measurements, our simulations framework was able to predict
diffuse reflections and transmissions in multi-layered translucent objects composed of base material
layers.
In the proposed research, we will extend these methods to cover the complete appearance of
volumetric materials that accounts for both directly reflected light and light scattering within the
material volume. 
We will use efficient simulation methods that allows us to simulate the appearance
of complex structures composed from these base materials. We developed an initial framework for
simulating behavior of multi-layered materials, where each layer is a homogeneous base material.
Then we extend this simulation to materials in which each voxel is made of a different measured
material. Since this general simulation is expensive and requires path-tracing algorithms, we used
algorithms that are based on fast numerical integration of our low-dimensional representations
and the use of pre-computation, similar to pre-computed light transport approaches \cite{Sloan:2002}.
\subsection{Simulation of Deformation Behavior}
In previous work, the problem
of modeling and simulating soft materials and their deformation behavior \cite{Bickel:2010,Bickel:2009}.
In order to accurately model deformation behaviors they built a system that allowed measuring how
a given material deforms subject to a variety of external forces. 
The system used an automated
robot arm to apply desired forces and a vision subsystem to measure corresponding surface deformations.
We modeled deformation behavior of a material using a stress-strain relationship. Since
the deformation behavior of many of our materials was non-linear, we developed a model that could
accurately represent this behavior and that was also easy to fit to the surface measurements. In
our model, the stress-strain relationship was represented as a collection of linear stress-strain relationships.
We used a properly chosen function to interpolate between these linear relationships
and obtain a non-linear model. Using this system and model we could accurately measure material
deformation of a collection of materials produced by a multi-material printer and fit these observation
to the stress-strain model. Then, using this material model within a Finite Element Method
we could accurately simulate behavior of any shape made from a composite of measured materials.
In the proposed research, we would like to address the shortcomings of our current measurement
and models. First, our current approach handles only elastic material deformation. We would like to
extend our current data-driven model to viscoelastic materials. In fact, most real materials exhibit
this property, which is characterized by hysteresis (or material memory). For these materials, it is
not possible to predict the behavior of a material only based on the currently applied stress. One also
has to know the history (or the sequence) of applied stress values to accurately simulate material
behavior. In order to determine the material model one has to measure the so called hysteresis
loops in the stress-strain space (as opposed to a single stress-strain curve for elastic materials). The
hysteresis loop is the result of the damping effect which causes the material to behave differently
depending whether stress is applied or removed as some of the energy is dissipated within the
material into heat. Additional advantage of using this material model is that it will allow us to
handle dynamic material deformations.

In order to measure viscoelastic material properties, we will need to modify our current measurement
system and procedures. We will use repeatable, time-coded measurements to estimate
Figure 5: From left to right: Force-and-deformation capture of a non-linear heterogeneous pillow;
synthesized deformation with fitted material parameters; and interactive deformation synthesized
with our soft tissue modeling technique using scattered data interpolation \cite{Bickel:2010}.

We will also use faster resonating actuators and high-speed
camera systems to track material deformations at high frame rates. Furthermore, we plan to track
the whole material volume rather than the outer material surface. This is possible since we can
print materials in clear color as well. Thus small color-coded markers can be easily embedded into
the material structure. These markers can be tracked in 3D using a multi-camera system.
\subsection{Optimization for Fabrication}
\cite{Finckh:2010} computing surface height field
of transparent material to match caustics pattern.
ShadowPix \cite{Bermano:2012} produced surface geometry
that produces different images 
under different lighting
directions. \cite{sko:2012} produced balloons that match 3D models when
inflated.
\subsection{Shade Trees}
\note{Should talk about influences from Shade Trees and Shader networks, note that these methods do not have tunable parameters}
\section{Contributions}
The overall contribution of our work is a flexible framework for the construction of material assignment algorithms. Our second contribution is the ``Reducer Network'',  a graphical data structure which describes both the geometric partitioning and material parameterization of  a given object. We couple this with a ``Tuner Network'' which automatically tunes the parameters of the Reducer Network to produce a fabricable object. Finally we will describe our API which utilizes these structures to produce numerous material assignment algorithms from a set of primitive operations.
\section{Methods}
Below we will describe the key components of our framework and its associated API. 
\subsection{The Reducer Network}
Computing material assignments at output device resolution is computationally intractable. Because of this, material assignments are computed  using a reduced representation. We specify this representation using a Reducer Network.  This structure is conceptually similar to those used in programmable shading systems (e.g. ~\cite{Cook1984}, Maya Shader Networks) . However these systems are primarily concerned with assigning known materials and textures to an object's surface whereas we are seeking a volumetric assignment of (as yet) unknown materials. In order to accomplish this task we build a directed, acyclic graph which has the object's surface geometry as its root node. We define two classes of nodes, geometry nodes and parameterization nodes. A geometry nodes takes a volumetric region as input and produces a partition of unity of this region as output. The leaf nodes of our graph are the parameterization nodes which associate a material assignment function of the form $\lambda\left(x,y,z,\theta\right)$ to an input volumetric region. Lambda can either output material parameters or a discrete material assignment for a spatial point $\left(x,y,z\right)$ given some parameters, here denoted by $\theta$. Because the network describes a nested set of partitions of unity we can efficiently perform material queries by passing a query point through the network until it arrives at a parameterization node. We note that Reducer Networks do not inherently enforce material continuity between disjoint regions in the graph however, for material assignment problems this is not required.

 See \textbf{Figure 1} for an example of a Reducer Network. \note{Desai: Make figure 1}.

\subsection{Tuners}
\note{Desai: can you write this section, I don't know the details}
\subsection{Tuning the Reducer Network}

\subsection{Tuner Network}
\note{Desai: also this section :) }
%\section{System Description}
%The software framework (shown in Figure~\ref{fig:framework}) is composed of 
%a set of software
%modules that can be easily modified or exchanged. The first component
%of this framework is an output-device-independent description API
%that allows the user to specify any physical object in terms of its physical
%properties such as geometry, mechanical behaviors, and appearance. 
%The framework will also provide physically-based
%simulation tools that accurately predict the behavior of our output device.
%The goal of the optimization module is to search through
%the space of possible device outputs and deliver one that best matches the
%desired model. In order to evaluate the quality of the match between the
%simulation output and the desired model we need to develop appropriate
%comparison metrics. Finally, a central component of our framework is
%a reduction model which makes the optimization problem tractable by
%reducing the search space as much as possible.
%The reduction model can specified by experts or deduced from a
%set of examples. Next, we will describe each element of the framework in
%more detail.
%\begin{figure}
%\includegraphics[width=0.45\textwidth]{figure/framework.pdf}
%\caption{API overview. The diagram will become nicer.}
%\label{fig:framework}
%\end{figure}
%\subsection{Input API}
%Our input API handles three tasks: specifying an optimization problem,
%maintaining a material library and configuring existing modules of
%the optimization process.
%An overview of the three components of our API is shown in Figure~\ref{fig:inputAPI}
%
%\begin{figure}
%\includegraphics[width=0.45\textwidth]{figure/inputAPI.pdf}
%\caption{Input API. The diagram will become nicer.}
%\label{fig:inputAPI}
%\end{figure}
%
%Our input API allows a designer to either
%define material properties for sections of an object
%or specify behavior of an object.
%The designer
%can use functions to describe properties such as reflectance (e.g., BRDFs, BSSRDFs), 
%stiffness, or other spatially varying properties. 
%Although it is possible to define the object directly by only specifying material
%properties, intuitive user
%interfaces will aid in specifying the objects and their properties
%in terms of intuitive metaphors.
%For example, the elastic properties of an object can be specified by a set of
%example displacements from the rest pose and the corresponding applied forces.
%\subsection{Search Space Decomposition and Reduction}
%Our framework includes an optimization component that
%searches through the space of possible device inputs
%to find the one that best reproduces
%the desired object. Unfortunately, this search space
%is usually intractable. For example, when the printing
%volume has N voxels and each of these voxels can be
%assigned to one of M base materials, the search space
%contains $N^M$ choices. 
%To overcome this problem, we decompose the search
%space into similar subproblems and also reduce
%the search space to a lower-dimensional space using a reduction model. 
%The goal of the reduction step is to
%shrink the search space as much as possible but at the
%same time not to prune out good solutions. The low-dimensional space produced by the
%reducer can be defined in two different ways. First, a reduction model can be explicitly specified by an expert, for
%example, by defining structures of base materials and rules how different base materials can be combined together.
%In this paper, all reduction models are manually specified by us.
%The second way to specify a reduction model is implicit. In this case, the user provides a set of examples of
%valid material structures. Then, a machine learning algorithm infers both basic material structures and the rules
%for combining them. The optimization framework can be further improved by employing techniques that cluster
%partial solutions that yield similar output properties, avoiding combinatorial explosion of the search space.
%Furthermore, we can employ bounds on the solution based on physical constraints of the base materials.
%\subsection{Optimization}
%Our optimization component is an interface to
%existing standard optimization techniques
%such as simulated annealing, variational inference and
%Markov Chain Monte Carlo methods,
%as well as custom optimization algorithms implemented
%by more advanced users.
%\subsection{Simulation}
%Another key step in the process of converting abstractions to physical output is being able to
%simulate what an output device is going to generate given a well-defined input. 
%A simulation allows the user to preview how the output will look and behave. 
%This simulation is a generalization of the print-preview function
%available in word processing applications.
%As physical output generation might be costly or time-consuming, 
%it is extremely beneficial for users to be able to preview the output and make immediate changes to the design.
%That means that the simulation must accurately predict the output. 
%It's necessary to develop very efficient and accurate rendering
%and finite element simulation packages. Since we run the simulation multiple times within the optimization method
%we can cache and reuse partial simulation results, speeding up the evaluation process at least 10-100 times.
%Furthermore, we will ensure that the simulation is accurate by measuring properties of the printing materials and using
%these measurements to estimate the parameters of data-driven material models (e.g., for elasticity, reflectance).
%
%\section{Built-in Components}
%We provide a built-in library that handles common tasks such as assigning texture,
%simulating deformation, and optimizing for a stacking.
%\subsection{Generator}
%Generator uses a reduced representation
%of material assignments.
%Each generator is defined over a domain in the input shape.
%We provide a few basic generators and a composition operator.
%Let the input shape be over a domain $\Omega$.
%The composition of two generators $G_2\circ G_1$ is implemented by
%first letting $G_1$ reduce over a domain $\Omega_1$ and
%then letting $G_2$ reduce $\Omega_2 = \Omega-\Omega_1$.
%An example is shown in Figure~\ref{fig:genCompose}
%\begin{figure}[h]
%\includegraphics[scale=0.2]{figure/genCompose.png}
%\caption{Composition of two generators. $G_1$ is a fine-grained
%surface voxelization and $G_2$ is a very coarse inner
%grid.}
%\label{fig:genCompose}
%\end{figure}
%
%We will provide the following generators.
%\begin{enumerate}
%\item A generator that uses a uniform grids with homogeneous
%material in each grid cell. (A foam structure is treated as one homogeneous material).
%\item A generator that decomposes a mesh into columns of stackings.
%\item A polynomial spline for outline of a material.
%\end{enumerate}
%We use composition of the above basic generators to make the examples in 
%the result section.
%\subsection{Simulator}
%\begin{enumerate}
%\item Physics simulator with FEM, rigid body simulation
%and simple collision resolution.
%\item nVidia optiX to do photon mapping for caustics. Maybe also BRDF and shadows.
%We need to look at its speed.
%\item local computations such as reflectance and local mechanical properties given a stacking.
%\end{enumerate}
%
%\subsection{Metric}
%Optionally returns a gradient represented as an array of signed floats.
%\begin{enumerate}
%\item Compare a mesh and its deformed state. There is one-to-one correspondence of 
%vertices.
%\item compare two images
%\item compare two real numbers such as weight and bouncing height.
%\end{enumerate}
%
%\subsection{Optimizer}
%\begin{enumerate}
%\item MCMC, Simulated Anealing, SPSA
%\item Branch and bound (plus clustering of similar configurations)
%\item (Optional) reinforcement learning, deep neural networks...
%\end{enumerate}
%
%\section{Programming with Our API}
%A casual user mostly make function calls to an existing 
%manager to pass data and to run the optimization.
%More advanced user can configure the manager to achieve more specific requirements.
%For example, he can swap out an existing optimization module to achieve a better accuracy or 
%a faster speed.
%A professional user can implement more modules when he wants to deal with new
%material properties that are not handled by our built-in library. In that case,
%he needs to know what value the new property is in print materials, he needs to know
%how to simulate for the new property, and how to measure error of the new property.
%
%In below, we provide code snippets that handles three common tasks
%that we expect a moderately advanced user would write.
%
%Code examples:
%To generate textured model:
%\begin{verbatim}
%Manager manager;
%manager.loadMaterialLib("material_file");
%manager.setGenerator(
%  new HeightFieldGenerator());
%manager.setSimulator(colorSimulator());
%manager.setMetric(someColorMetric());
%manager.setOptimizer(
%  simulatedAnealing);
%  //or e.g. gradientDescent
%Objective obj;
%Scene scene;
%TexturedShape shape("filename");
%TexturedShape goal("filename1");
%scene.addShape(shape);
%obj.setScene(scene);
%obj.setGoal(goal);
%manager.addObjective(obj);
%\end{verbatim}
%
%To optimize BRDF for a particular camera view point and lighting.
%\begin{verbatim}
%scene.setCamera(camera);
%scene.addLight(light);
%manager.setGenerator(stackingGenerator);
%manager.setSimulator(brdfSimulator);
%manager.setMetric(someImageMetric);
%Shape shape("filename"), 
%Image goal("filename1");
%\end{verbatim}
%
%Defining elasticity properties
%by specifying deformation 
%reaction to a force applied to a point.
%
%\begin{verbatim}
%force = new Force(pos, dir, shapeID);
%constraint = new Constraint(
%    FIX_VERTEX, vertexList, shapeID));
%
%scene.addForce(force);
%scene.addConstraint(constraint);
%Shape goal;
%manager.setSimulator(femSimulator);
%\end{verbatim}

\section{Results}
We used various generators in our experiments.
\subsection{Print Material Characterization and Measurement}
\subsection{Subsurface Scattering, Deformation and Mass}
Combining previous work on replicating
subsurface scattering and deformation behavior, 
we produced
material with a particular
look and deformation properties as shown in Figure~\ref{fig:deform}.
We als considered one additional property:mass. 
In this example, we
tried to make the material as light as possible.

We tried different combinations of generators in this experiment.
We compared the performance of these different generator
in terms of run-time and error.
\begin{enumerate}
\item Spline for top layer, fine-grained voxelization for mid layer
and a coarse grid for bottom layer.
\item Voxels for top layer, splines for middle layer and a coarse
bottom layer.
\item Find-grained voxels for top layer and coarse grid for bottom layer.
\end{enumerate}

\begin{figure}
\begin{subfigure}[b]{0.3\textwidth}
	\centering
 	\includegraphics[width=\textwidth]{figure/slab_origin.png}
   	\caption{}
    \label{fig:deformPhoto}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.3\textwidth}
    \centering
  	\includegraphics[width=\textwidth]{figure/slab_deform.png}
    \caption{}
    \label{fig:deformPrint}
\end{subfigure}
\caption{Simutaneously matching appearance and deformation. 
(a)Original slab of a particular looking. (b)printed slab 
with the same looking and desired deformation.}
\label{fig:deform}
\end{figure}

Two error plots showing subsurface scattering and deformation errors decrease
after each iteration in Figure~\ref{fig:err}.

\begin{figure}
\begin{subfigure}[b]{0.3\textwidth}
	\centering
 	\includegraphics[width=\textwidth]{figure/error.pdf}
   	\caption{}
    \label{fig:ssErr}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.3\textwidth}
    \centering
  	\includegraphics[width=\textwidth]{figure/error.pdf}
    \caption{}
    \label{fig:deformErr}
\end{subfigure}
\caption{Error plots for 
optimizing subsurface-scattering and deformation properties 
iteratively. (a)Subsurface scattering errors. (b)Deformation errors.}
\label{fig:err}
\end{figure}

\subsection{Caustics for a Mesh}
We demonstrate how to produce a 3D model with specific caustic shape under a particular lighting,
similar to \cite{Finckh:2010}. The difference is that we optimize on an arbitrary 3D model,
and we use a generic rendering package (NVidia optiX). 
We represented the surface with a spline surface.
Our algorithm took 6 hours to finish
with $4000$ spline control points for the model and $600\times 600$ resolution for the caustic image. A 
picture of our printed result under desired lighting is shown in Figure~\ref{fig:caustic}.

\begin{figure}
	\centering
 	\includegraphics[width=0.3\textwidth]{figure/caustic.jpg}
\caption{Teapot with a caustic image.}
\label{fig:caustic}
\end{figure}

\subsection{Textured Model}
Input: a textured 3D model and
measured Albedo of print material.

Output: (Probably gray-scale) Material arranged for different printers in different formats: STL, fable,Gcode.

Printers: Our printer, Makerbot, Objet, maybe Zcorp.

When printing texture, only the outer layer matters. We used 
a composition of two generators: a very high resolution grid
for the outer layer and a uniform material for the inside.

We expand the set of printable colors by stacking translucent materials. We then applied
dithering with our expanded colors.
Results are shown in Figure~\ref{fig:texture}
\subsection{New Examples for Mechanical Properties}
Ball bounce to certain height. We used a higher resolution
grid for the out layer and very low resolution
for the inner mesh.

A plot of bounce height after each iteration.
The height and desired height is plotted in Figure~\ref{fig:ball}
\begin{figure}
	\centering
 	\includegraphics[width=0.3\textwidth]{figure/ballHeight.pdf}
\caption{Simulated height per-iteration of a ball bouncing after free-falling at 1 meter height.
	Redline is desired height.}
\label{fig:ball}
\end{figure}

A loaded die as shown in Figure~\ref{fig:die}. This example is interesting because
a local change will have some global effect.
We used very high resolution grid for the surface to match texture
of a die. The inner lay has only $3\times 3$ resolution.
\begin{figure}
	\centering
 	\includegraphics[width=0.25\textwidth]{figure/die.png}
\caption{A printed loaded die}
\label{fig:die}
\end{figure}
It has a certain probability to land on "1".
A table of probability distribution of landing on each side
is shown in Table~\ref{tab:dice}.

\begin{table}
\centering
  \begin{tabular}{ |c| p{0.7in} | p{0.7in} | }
  \hline
  face & simulated probability & measured probability\\
  \hline
  1 & 0.5 & 0.5\\
  \hline
  2 & 0.1 & 0.1\\
  \hline
  3 & 0.1 & 0.1\\
  \hline
  4 & 0.1 & 0.1\\
  \hline
  5 & 0.1 & 0.1\\
  \hline
  6 & 0.1 & 0.1\\

  \hline
  \end{tabular}
  \caption{Simulated and measured dice probability}
  \label{tab:dice}
\end{table}

A plot of simulated probability of landing on "1" after each iteration
is shown in Figure~\ref{fig:dieErr}.

\begin{figure}
	\centering
 	\includegraphics[width=0.3\textwidth]{figure/die.pdf}
\caption{Probability of die landing on 1 after each iteration.}
\label{fig:dieErr}
\end{figure}

Printer:objet, ours.

More ideas:
balancing bird with texture.
\section{Possible Extension}
UI to specify input deformation.
\section{Conclusion}
\section*{Acknowledgements}

\bibliographystyle{acmsiggraph}
\bibliography{paper}
\end{document}