### 2013-05-03 thesis/1_cover.tex
add tex files

%%
%% MIT Thesis - Title Page
%%

\thispagestyle{empty}

\begin{centering}
\rule{\textwidth}{.05in}\\ \vspace{.10in}

\Large \textbf{\textsf{ Spec2Fab: A Reducer-Tuner Model for Translating Specifications to 3D Prints
}} \\
\normalsize

\vspace{.1in}
by \\

\vspace{.1in}
Donglai Wei \\

\vspace{.1in} B.S., Computer Science,
Carnegie Mellon University, 2011 \\

\vspace{.0in} S.M., Electrical Engineering and Computer Science,
M.I.T., 2013 \\

\vspace{.1in}
\rule{\textwidth}{.05in}\\

\vspace{.2in}

Proposal for Thesis Research in Partial Fuillment\\
of the requirements for the degree of Master of Science\\

\vspace{.15in}

\vspace{.1in} March 2015

\vspace{.15in} \copyright\ 2015
Massachusetts Institute of Technology \\
All Rights Reserved. \\

\end{centering}

\vspace{.3in}

\noindent
Signature of Author: \underline{\mbox{\hspace{4.28in}}} \\
\vspace{-.25in}
{\flushright Department of Electrical Engineering and Computer Science \\
March 18, 2015\\
}

\vspace{.15in}

\noindent
Certified by: \underline{\mbox{\hspace{4.827in}}} \\
\vspace{-.25in}
{\flushright Your first supervisor's name here \\
Title of supervisor, e.g., Professor of Electrical Engineering and Computer Science \\
Thesis Supervisor \\
}

\vspace{.15in}

\noindent
Certified by: \underline{\mbox{\hspace{4.827in}}} \\
\vspace{-.25in}
{\flushright Your second supervisor's name here \\
Title of supervisor, e.g., Professor of Electrical Engineering and Computer Science \\
Thesis Supervisor \\
}

\vspace{.15in}

\noindent
Accepted by: \underline{\mbox{\hspace{4.79in}}} \\
\vspace{-.25in}
{\flushright Leslie A.~Kolodziejski \\
Professor of Electrical Engineering and Computer Science \\
Chair, Committee for Graduate Students \\
}

\cleardoublepage