### 2013-05-03 thesis/1_cover_thesis.tex

%%
%% MIT Thesis - Title Page
%%

\thispagestyle{empty}

\begin{centering}
\rule{\textwidth}{.05in}\\ \vspace{.10in}

\Large \textbf{\textsf{ Your title goes here
}} \\
\normalsize

\vspace{.1in}
by \\

\vspace{.1in}
Your name goes here \\

\vspace{.1in} B.S., Electrical Engineering,
some other school, 2010 \\

\vspace{.0in} S.M., Electrical Engineering and Computer Science,
M.I.T., 2012 \\

\vspace{.1in}
\rule{\textwidth}{.05in}\\

\vspace{.2in}

Submitted to the Department of Electrical Engineering and Computer Science \\
in partial fulfillment of the requirements for the degree of \\

\vspace{.15in}

Doctor of Philosophy \\ % <--- replace this with Master of Science if you want a master's instead
in Electrical Engineering and Computer Science \\
at the Massachusetts Institute of Technology \\

\vspace{.1in} March 2015

\vspace{.15in} \copyright\ 2015
Massachusetts Institute of Technology \\
All Rights Reserved. \\

\end{centering}

\vspace{.3in}

\noindent
Signature of Author: \underline{\mbox{\hspace{4.28in}}} \\
\vspace{-.25in}
{\flushright Department of Electrical Engineering and Computer Science \\
March 18, 2015\\
}

\vspace{.15in}

\noindent
Certified by: \underline{\mbox{\hspace{4.827in}}} \\
\vspace{-.25in}
{\flushright Your first supervisor's name here \\
Title of supervisor, e.g., Professor of Electrical Engineering and Computer Science \\
Thesis Supervisor \\
}

\vspace{.15in}

\noindent
Certified by: \underline{\mbox{\hspace{4.827in}}} \\
\vspace{-.25in}
{\flushright Your second supervisor's name here \\
Title of supervisor, e.g., Professor of Electrical Engineering and Computer Science \\
Thesis Supervisor \\
}

\vspace{.15in}

\noindent
Accepted by: \underline{\mbox{\hspace{4.79in}}} \\
\vspace{-.25in}
{\flushright Leslie A.~Kolodziejski \\
Professor of Electrical Engineering and Computer Science \\
Chair, Committee for Graduate Students \\
}

\cleardoublepage